Leadership across cultures and contexts: an exploratory analysis

### 2020-04-12 leadership_across_cultures_contexts.Rmd
Minor edit. Changed a subheading.

We then explored "features" in our data. There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. We used several methods to explore lower dimensional representations of these data that might reveal informative structure, including logistic PCA, non-negative matrix factorization (NMF), and agglomerative hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA]. NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010]. Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

All analyses were conducted with `r R.version.string`.

@@

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.

@@

## Variable distributions

>>>>

We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA]. NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010]. Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

The analyses computing levels of evidence for each variable used the entire set of text records. Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros). All analyses were conducted with `r R.version.string`.

@@

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.

@@

## Evidence for leader functions and qualities
